---
title: "Physiology_stats"
author: "Melissa Naugle"
date: "5/10/2021"
output: html_document
---

#Load packages and data

```{r }
#change this to your working directory
setwd(dir = "~/Desktop/GitHub/CBASS_bleachingdata/")
rm( list = ls())
graphics.off()
library(ggplot2)
library(plyr)
library(dplyr)
library(reshape2)
library(mosaic)
library(car)
library(agricolae)
library(lmerTest)
#library(emmeans)
library(ggpubr)
library(tidyr)

data <- read.csv("raw data sheets/MasterASData.csv")
head(data)
data$Time.point <- as.character(data$Time.point)
#insert "field" into timepoint for field collected corals
data$Time.point[data$Ramp == "Field"] <- "Field"


```

This script goes through my three physiology measurements- Red intensity, PAM, and color card. For each measurement I am asking: 
1. Are there differences in (red intensity, PAM, and color card) in corals at diff temperature treatments? (4 temp treatments (28,32,33,34))
2. Are there differences in (red intensity, PAM, and color card) in corals from diff sites (5 sites)?

For all of these, I am normalizing values to those at control temperatures 



# Red Intensity

Recovery:

Assumptions met: shapiro test and levene test 

Now we run an ANOVA on the normalized data

Ran full model: interaction of site and ramp is not signif (p = 0.3948698)

Without interaction term: site signif (p = 0.02162), temp signif (p = 5.043e-05)

Variances equal and residuals normal 

Tukey posthoc: 
site: 
vatia-cannery 0.0473585
vatia-fagatele 0.0270875

temp: 
33 is diff than 34 0.0002497
33 is diff than 35 0.0004170

Stress: 

All assumptions met
Site not signif F = 0.3409, p = 0.7957, df = 3, 180
ramp not signif F = 0.2194, p = 0.8032, df = 2, 180 
interaction not signif F = 0.5244, p = 0.7893, df = 6, 180 

without interaction term 
Site not signif F = 0.7282, p = 0.5364, df = 3, 186
ramp not signif F = 2.2300, p = 0.1104, df = 2, 186

```{r }

#remove NAs
data <- data[!is.na(data$AverageRed), ]

#reorganize data to normalize to controls
data$Coral.Watch.Color1 <- NULL
data$Coral.Watch.Color2 <- NULL
data$Coral.Watch.Color3 <- NULL
data$PAM..Fv.Fm. <- NULL
data$X <- NULL
#if tank A-D, rep 1; if tank E-F, rep 2
data$Rep[data$Tank == "A"] <- "1"
data$Rep[data$Tank == "B"] <- "1"
data$Rep[data$Tank == "C"] <- "1"
data$Rep[data$Tank == "D"] <- "1"
data$Rep[data$Tank == "E"] <- "2"
data$Rep[data$Tank == "F"] <- "2"
data$Rep[data$Tank == "G"] <- "2"
data$Rep[data$Tank == "H"] <- "2"
data$Tank <- NULL
data_normalized <- spread(data, Ramp, AverageRed)
head(data_normalized)
data_normalized$t35_normalizedto28 <- data_normalized$`35` - data_normalized$`28`
data_normalized$t34_normalizedto28 <- data_normalized$`34` - data_normalized$`28`
data_normalized$t33_normalizedto28 <- data_normalized$`33` - data_normalized$`28`
head(data_normalized)

data_normalized_stress <- data_normalized %>%
  filter(data_normalized$Time.point == "Stress") 

data_normalized_recovery <- data_normalized %>%
  filter(data_normalized$Time.point == "Recovery") 

write.csv(data_normalized_recovery, "clean data/avgred_data_normalized_recovery.csv")

#plot data
data_normalized_stress <- gather(data_normalized_stress, Ramp, AverageRed, t35_normalizedto28:t33_normalizedto28)
stress <- ggplot(data=data_normalized_stress, aes(x=Site, y=AverageRed, fill = Ramp))+ geom_boxplot() + 
  theme_bw() + ylab("Red Intensity \n (normalized to controls at 28C)") + xlab("") + scale_fill_discrete(name = "Treatment", labels = c("33C", "34C", "35C"))  + scale_fill_manual(values = c("dodgerblue4", "springgreen3", "yellow2"), name = "Treatment", labels = c("33C", "34C", "35C")) + ggtitle("Heat Stress") + theme(plot.title = element_text(hjust = 0.5))

data_normalized_recovery <- gather(data_normalized_recovery, Ramp, AverageRed, t35_normalizedto28:t33_normalizedto28)
recovery <- ggplot(data=data_normalized_recovery, aes(x=Site, y=AverageRed, fill = Ramp))+ geom_boxplot() + theme_bw() + ylab("Red Intensity \n (normalized to controls at 28C)") + xlab("") + scale_fill_discrete(name = "Treatment", labels = c("33C", "34C", "35C"))  + scale_fill_manual(values = c("dodgerblue4", "springgreen3", "yellow2"), name = "Treatment", labels = c("33C", "34C", "35C")) + ggtitle("Recovery") + theme(plot.title = element_text(hjust = 0.5))

ggarrange(stress, recovery, ncol = 1, nrow = 2, common.legend = TRUE, legend = "right")

#ggsave("t2_red_normalized.pdf", width = 6, height = 4)


```
All assumptions passed!

```{r assumptions}
#check assumptions
#stress

hist(data_normalized_stress$AverageRed) #looks normal!
ggdensity(data_normalized_stress$AverageRed) #looks normal
ggqqplot(data_normalized_stress$AverageRed) #looks normal!
shapiro.test(data_normalized_stress$AverageRed) #normal!!!!

#equality of variances
bartlett.test(AverageRed ~ Site, data=data_normalized_stress) #equal
bartlett.test(AverageRed ~ Ramp, data=data_normalized_stress) #equal
leveneTest(AverageRed ~ Site * Ramp, data=data_normalized_stress) #equal with both

#recovery 

hist(data_normalized_recovery$AverageRed) #looks normal!
ggdensity(data_normalized_recovery$AverageRed) #looks normal
ggqqplot(data_normalized_recovery$AverageRed) #looks normal!
shapiro.test(data_normalized_recovery$AverageRed) #normal!!!!

#equality of variances
bartlett.test(AverageRed ~ Site, data=data_normalized_recovery) #not equal (but that's okay)
bartlett.test(AverageRed ~ Ramp, data=data_normalized_recovery) #equal
leveneTest(AverageRed ~ Site * Ramp, data=data_normalized_recovery) #equal with both

#write.csv(data_normalized, "~/Desktop/GitHub/CBASS_bleachingdata/raw data sheets/AverageRedData_normalized.csv", row.names = F)
```



```{r ANOVA_redintensity}

#recovery

# Analysis of variance (2-way) ANOVA CBASS
red.CBASS.lm <- lm(AverageRed ~ Site * Ramp, data = data_normalized_recovery)
Anova(red.CBASS.lm, type = "III")
#site is not signif (p = 0.7818383)
#temp is signif (p = 0.0004361)
#interaction of site and ramp is not signif (p = 0.3948698)

#since the interaction is not signif, we can run ANOVA without interaction
red.CBASS.lm <- lm(AverageRed ~ Site + Ramp, data = data_normalized_recovery)
Anova(red.CBASS.lm, type = "III")
#site signif (p = 0.02162)
#temp signif (p = 5.043e-05)

# Model fitting and assumptions diagnostic - EOV
plot(AverageRed ~ interaction(Site,Ramp), data = data_normalized_recovery) # Box-plot homogeneity of variance - looks good
leveneTest(AverageRed ~ Site * Ramp, data=data_normalized_recovery) 
# equal (p = 0.104)

#plot residuals - -check normality 
plot(red.CBASS.lm, 1) # Residual vs Fitted values
qqnorm(resid(red.CBASS.lm)); qqline(resid(red.CBASS.lm)) #mostly normal 
hist(resid(red.CBASS.lm)) #looks perfectly normal
shapiro.test(red.CBASS.lm$residuals) #very normal - p = 0.7882


# try tukey post hoc 
tukey <- TukeyHSD(red.CBASS.lm)
plot(tukey)
tukey
HSD.test(red.CBASS.lm, trt = "Site", console = T)
#shows signif diffs for: 
#site: vatia-cannery and vatia-fagatele
#temp: 33 is diff than 34 and 35

#stress

# Analysis of variance (2-way) ANOVA CBASS
red.CBASS.lm <- lm(AverageRed ~ Site * Ramp, data = data_normalized_stress)
Anova(red.CBASS.lm, type = "III")

#since the interaction is not signif, we can run ANOVA without interaction
red.CBASS.lm <- lm(AverageRed ~ Site + Ramp, data = data_normalized_stress)
Anova(red.CBASS.lm, type = "III")

# Model fitting and assumptions diagnostic - EOV
plot(AverageRed ~ interaction(Site,Ramp), data = data_normalized_stress) # Box-plot homogeneity of variance - looks good
leveneTest(AverageRed ~ Site * Ramp, data=data_normalized_stress) # equal (p = 0.104)

#plot residuals - -check normality 
plot(red.CBASS.lm, 1) # Residual vs Fitted values
qqnorm(resid(red.CBASS.lm)); qqline(resid(red.CBASS.lm)) #mostly normal 
hist(resid(red.CBASS.lm)) #looks perfectly normal
shapiro.test(red.CBASS.lm$residuals) #very normal - p = 0.7882


# try tukey post hoc 
tukey <- TukeyHSD(red.CBASS.lm)
plot(tukey)
tukey
HSD.test(red.CBASS.lm, trt = "Site", console = T)
#shows signif diffs for: 
```


#PAM

Removed 4 outliers where PAM = 0 

Assumptions met: levene test and shapiro test (shapiro test not met, but but large sample size and balanced design, so ANOVA is robust) 

Now we run an ANOVA on the normalized data

Ran full model: interaction of site and ramp is not signif (p = 0.9242)

Without interaction term: site signif (p = 0.01626), temp not signif (p = 0.16714)

Variances equal and residuals somewhat normal (a bit skewed, but large sample size and balanced design, so ANOVA is robust)

Tukey posthoc: 
site: 
fagatele-fagaalu 0.0274991

```{r }

data <- read.csv("raw data sheets/MasterASData.csv")
data$Time.point <- as.character(data$Time.point)
#insert "field" into timepoint for field collected corals
data$Time.point[data$Ramp == "Field"] <- "Field"

data$Coral.Watch.Color1 <- NULL
data$Coral.Watch.Color2 <- NULL
data$Coral.Watch.Color3 <- NULL
data$AverageRed <- NULL
data$X <- NULL
#if tank A-D, rep 1; if tank E-F, rep 2
data$Rep[data$Tank == "A"] <- "1"
data$Rep[data$Tank == "B"] <- "1"
data$Rep[data$Tank == "C"] <- "1"
data$Rep[data$Tank == "D"] <- "1"
data$Rep[data$Tank == "E"] <- "2"
data$Rep[data$Tank == "F"] <- "2"
data$Rep[data$Tank == "G"] <- "2"
data$Rep[data$Tank == "H"] <- "2"
data$Tank <- NULL
data_normalized <- spread(data, Ramp, PAM..Fv.Fm.)
head(data_normalized)
data_normalized$t35_normalizedto28 <- data_normalized$`35` - data_normalized$`28`
data_normalized$t34_normalizedto28 <- data_normalized$`34` - data_normalized$`28`
data_normalized$t33_normalizedto28 <- data_normalized$`33` - data_normalized$`28`
head(data_normalized) #missing values for field samples and for stress for rep 2
#remove nas 
data_norm <- data_normalized[complete.cases(data_normalized[ , 13:15]),]

#write.csv(data_norm, "~/Desktop/GitHub/CBASS_bleachingdata/raw data sheets/PAM_data_normalized.csv", row.names = F)

#plot data
data_norm2 <- gather(data_norm, Ramp, PAM..Fv.Fm., t35_normalizedto28:t33_normalizedto28)
ggplot(data=data_norm2, aes(x=Site, y=PAM..Fv.Fm., fill = Ramp))+ geom_boxplot() + 
  theme_bw() + ylab("Red Intensity \n (normalized to controls at 28C)") + xlab("") + scale_fill_discrete(name = "Treatment", labels = c("33C", "34C", "35C"))  + scale_fill_manual(values = c("dodgerblue4", "springgreen3", "yellow2"), name = "Treatment", labels = c("33C", "34C", "35C"))

#remove those 4 outliers, where pam value = 0
nrow(data_norm2)
data_norm2 <- data_norm2[!(data_norm2$PAM..Fv.Fm. < -0.5),]
nrow(data_norm2)
#removed 4 rows 


#plot data
ggplot(data=data_norm2, aes(x=Site, y=PAM..Fv.Fm., fill = Ramp))+ geom_boxplot() + 
  theme_bw() + ylab("Red Intensity \n (normalized to controls at 28C)") + xlab("") + scale_fill_discrete(name = "Treatment", labels = c("33C", "34C", "35C"))  + scale_fill_manual(values = c("dodgerblue4", "springgreen3", "yellow2"), name = "Treatment", labels = c("33C", "34C", "35C"))
```


```{r assumptions}
#check assumptions
#normality
hist(data_norm2$PAM..Fv.Fm.) #looks normal!
ggdensity(data_norm2$PAM..Fv.Fm.) #looks normal
ggqqplot(data_norm2$PAM..Fv.Fm.) #looks mostly normal!
shapiro.test(data_norm2$PAM..Fv.Fm.) #not normal but kinda close!! p = 0.02398 so that's okay bc large sample size, balanced design 

#equality of variances
bartlett.test(PAM..Fv.Fm. ~ Site, data=data_norm2)
leveneTest(PAM..Fv.Fm. ~ Site, data=data_norm2) #equal
```

Run ANOVA on normalized PAM data 


```{r ANOVA_PAM}

# Analysis of variance (2-way) ANOVA CBASS
pam.CBASS.lm <- lm(PAM..Fv.Fm. ~ Site * Ramp, data = data_norm2)
Anova(pam.CBASS.lm, type = "III")
#site is not signif 
#temp is not signif 
#interaction of site and ramp is not signif (p = 0.9242)

#since the interaction is not signif, we can run ANOVA without interaction
pam.CBASS.lm <- lm(PAM..Fv.Fm. ~ Site + Ramp, data = data_norm2)
Anova(pam.CBASS.lm, type = "III")
#site signif (p = 0.01626)
#temp not signif (weird)

# Model fitting and assumptions diagnostic 
plot(PAM..Fv.Fm. ~ interaction(Site,Ramp), data = data_norm2) # Box-plot homogeneity of variance
leveneTest(PAM..Fv.Fm. ~ Site * Ramp, data=data_norm2) # equal 

#plot residuals 
plot(pam.CBASS.lm, 1) # Residual vs Fitted values
qqnorm(resid(pam.CBASS.lm)); qqline(resid(pam.CBASS.lm)) #mostly normal 
hist(resid(pam.CBASS.lm)) #looks mostly normal
shapiro.test(pam.CBASS.lm$residuals) #not normal, but close enough 

# try tukey post hoc 
tukey <- TukeyHSD(pam.CBASS.lm)
plot(tukey)
tukey
#shows signif diffs for: 
#site: fagatele-fagaalu

HSD.test(pam.CBASS.lm, trt = "Site", console = T)


```

# Color Card

Assumptions met: levene test and shapiro test (shapiro test not met (p = 0.001012), but but large sample size and balanced design, so ANOVA is robust) 

Now we run an ANOVA on the normalized data

Ran full model: interaction of site and ramp is  signif (F = 2.2153, p = 0.027332, df = 8,222)

Without interaction term: site not signif (F = 2.1798 ,p = 0.072187, df = 4,222), temp signif (F = 13.3836, p = 3.252e-06, df = 2, 222)

Variances equal and residuals normal 

Tukey posthoc: 
site: 
Faga'alu-Cannery 0.0331321
Vatia-Coconut Point 0.0410378
Vatia-Faga'alu 0.0004524

temp:
33-34 0.0000000
33-35 0.0000000

interaction:
is signif!! 
Cannery:t34_normalizedto28-Cannery:t33_normalizedto28         1.543691e-04
Coconut Point:t34_normalizedto28-Cannery:t33_normalizedto28   9.672545e-04
Faga'alu:t34_normalizedto28-Cannery:t33_normalizedto28        4.633011e-05
Cannery:t35_normalizedto28-Cannery:t33_normalizedto28         1.449874e-02
Coconut Point:t35_normalizedto28-Cannery:t33_normalizedto28   8.739863e-05
Faga'alu:t35_normalizedto28-Cannery:t33_normalizedto28        3.600332e-06
Faga'tele:t35_normalizedto28-Cannery:t33_normalizedto28       9.374453e-07
Faga'alu:t35_normalizedto28-Coconut Point:t33_normalizedto28  1.336939e-02
Faga'tele:t35_normalizedto28-Coconut Point:t33_normalizedto28 5.271960e-03
Cannery:t34_normalizedto28-Faga'tele:t33_normalizedto28       1.449874e-02
Faga'alu:t34_normalizedto28-Faga'tele:t33_normalizedto28      5.662989e-03
Coconut Point:t35_normalizedto28-Faga'tele:t33_normalizedto28 8.734804e-03
Faga'alu:t35_normalizedto28-Faga'tele:t33_normalizedto28      7.063657e-04
Faga'tele:t35_normalizedto28-Faga'tele:t33_normalizedto28     2.278777e-04
Cannery:t34_normalizedto28-Vatia:t33_normalizedto28           1.951206e-02
Faga'alu:t34_normalizedto28-Vatia:t33_normalizedto28          7.809006e-03
Coconut Point:t35_normalizedto28-Vatia:t33_normalizedto28     1.185714e-02
Faga'alu:t35_normalizedto28-Vatia:t33_normalizedto28          1.016775e-03
Faga'tele:t35_normalizedto28-Vatia:t33_normalizedto28         3.343505e-04
Vatia:t35_normalizedto28-Faga'alu:t35_normalizedto28          4.510750e-02
Vatia:t35_normalizedto28-Faga'tele:t35_normalizedto28         1.951206e-02

```{r }

data <- read.csv("raw data sheets/MasterASData.csv")
data_cw <- data
head(data_cw)
data_cw$AverageRed <- NULL
data_cw$PAM..Fv.Fm. <- NULL
data_cw$X <- NULL
#if tank A-D, rep 1; if tank E-F, rep 2
data_cw$Rep[data_cw$Tank == "A"] <- "1"
data_cw$Rep[data_cw$Tank == "B"] <- "1"
data_cw$Rep[data_cw$Tank == "C"] <- "1"
data_cw$Rep[data_cw$Tank == "D"] <- "1"
data_cw$Rep[data_cw$Tank == "E"] <- "2"
data_cw$Rep[data_cw$Tank == "F"] <- "2"
data_cw$Rep[data_cw$Tank == "G"] <- "2"
data_cw$Rep[data_cw$Tank == "H"] <- "2"
data_cw$Tank <- NULL
data_cw
head(data_cw)
data_cw_1 <- data_cw
data_cw_2 <- data_cw
data_cw_3 <- data_cw
data_cw_1$Coral.Watch.Color2 <- NULL
data_cw_1$Coral.Watch.Color3 <- NULL
data_cw_2$Coral.Watch.Color1 <- NULL
data_cw_2$Coral.Watch.Color3 <- NULL
data_cw_3$Coral.Watch.Color1 <- NULL
data_cw_3$Coral.Watch.Color2 <- NULL

data_normalized_CW1 <- spread(data_cw_1, Ramp, Coral.Watch.Color1)
head(data_normalized_CW1)
data_normalized_CW2 <- spread(data_cw_2, Ramp, Coral.Watch.Color2)
head(data_normalized_CW2)
data_normalized_CW3 <- spread(data_cw_3, Ramp, Coral.Watch.Color3)
head(data_normalized_CW3)


data_normalized_CW2$t35_normalizedto28 <- data_normalized_CW2$`28` - data_normalized_CW2$`35`
data_normalized_CW2$t34_normalizedto28 <- data_normalized_CW2$`28` - data_normalized_CW2$`34`
data_normalized_CW2$t33_normalizedto28 <- data_normalized_CW2$`28` - data_normalized_CW2$`33`


data_normalized_CW3$t35_normalizedto28 <- data_normalized_CW3$`28` - data_normalized_CW3$`35`
data_normalized_CW3$t34_normalizedto28 <- data_normalized_CW3$`28` - data_normalized_CW3$`34`
data_normalized_CW3$t33_normalizedto28 <- data_normalized_CW3$`28` - data_normalized_CW3$`33`

#T3

data_normalized <- data_normalized_CW3

data_normalized$StressRecovTimePoint <- data_normalized$Time.point
data_normalized$Time.point <- NULL

data_normalized_gather <- pivot_longer(data_normalized, cols = starts_with("t"), names_to = "temperature_norm", values_to = "colorcard" )
data_normalized_gather <- data_normalized_gather[!is.na(data_normalized_gather$colorcard),]

#T2
data_normalized_t2 <- data_normalized_CW2

data_normalized_t2$StressRecovTimePoint <- data_normalized_t2$Time.point
data_normalized_t2$Time.point <- NULL

data_normalized_gather_t2 <- pivot_longer(data_normalized_t2, cols = starts_with("t"), names_to = "temperature_norm", values_to = "colorcard" )
data_normalized_gather_t2 <- data_normalized_gather_t2[!is.na(data_normalized_gather_t2$colorcard),]


```


```{r assumptions}
#check assumptions
#normality
hist(data_normalized_gather$colorcard) #looks normal!
ggdensity(data_normalized_gather$colorcard) #looks normal
ggqqplot(data_normalized_gather$colorcard) #looks mostly normal!
shapiro.test(data_normalized_gather$colorcard)
#not normal but kinda close!! p = 0.001012

#equality of variances
leveneTest(colorcard ~ Site, data=data_normalized_gather)
bartlett.test(colorcard ~ Site, data=data_normalized_gather)
#equal

#t2
hist(data_normalized_gather_t2$colorcard) #looks normalish!
shapiro.test(data_normalized_gather_t2$colorcard) #not normal

leveneTest(colorcard ~ Site, data=data_normalized_gather_t2) #notequal 


```


```{r ANOVA_colorcard}
# Analysis of variance (2-way) ANOVA CBASS
colorcard.lm <- lm(colorcard ~ Site * temperature_norm, data = data_normalized_gather)
Anova(colorcard.lm, type = "III")

#plot residuals 
plot(colorcard.lm, 1) # Residual vs Fitted values
qqnorm(resid(colorcard.lm)); qqline(resid(colorcard.lm)) #mostly normal 
hist(resid(colorcard.lm)) #looks mostly normal
shapiro.test(colorcard.lm$residuals) #normal

# try tukey post hoc 
tukey <- TukeyHSD(colorcard.lm)
plot(tukey)
tukey


subset(tukey$`Site:temperature_norm`, tukey$`Site:temperature_norm`[,4] < 0.05)
HSD.test(colorcard.lm, trt = "Site", console = T)


#t2
colorcard.lm <- lm(colorcard ~ Site * temperature_norm, data = data_normalized_gather_t2)
Anova(colorcard.lm, type = "III") #no interaction signif 

colorcard.lm <- lm(colorcard ~ Site + temperature_norm, data = data_normalized_gather_t2)
Anova(colorcard.lm, type = "III")


#plot residuals 
plot(colorcard.lm, 1) # Residual vs Fitted values
qqnorm(resid(colorcard.lm)); qqline(resid(colorcard.lm)) #mostly normal 
hist(resid(colorcard.lm)) #looks mostly normal
shapiro.test(colorcard.lm$residuals) #normal

# try tukey post hoc 
tukey <- TukeyHSD(colorcard.lm)
plot(tukey)
tukey


HSD.test(colorcard.lm, trt = "Site", console = T)

#Faga'tele bleached more than Cannery Coconut Point Faga'alu
#Cannery bleached more than coconut point
#Vatia bleached more than Coconut Point Faga'alu


#Faga'tele-Coconut Point
#Faga'tele-Cannery
#Faga'tele-Faga'alu
#Coconut Point-Cannery 
#Vatia-Coconut Point
#Vatia-Faga'alu

#tele D
#cannery BC
#vatia CD
#coconut point A
#alu AB




```



