---
title: "Physiology_figs"
author: "Melissa Naugle"
date: "10/23/2020"
output: html_document
---

```{r }
setwd(dir = "~/Desktop/GitHub/CBASS_bleachingdata/")
rm( list = ls())
graphics.off()
library(ggplot2)
require(plyr)
library(dplyr)
library(reshape2)
library(mosaic)
```


# Red Intensity 


Plot raw colorimetric data

Makes 4 plots:
1. controls (tank (28C) and field controls)
2. 35C
3. 34C
4. 33C

```{r avgreddata}
data <- read.csv("raw data sheets/MasterASData.csv")
head(data)
data$Time.point <- as.character(data$Time.point)
data$Time.point[data$Ramp == "Field"] <- "Field"

#filter out only controls
data_controls <- data %>%
  filter(data$Ramp == "28" | data$Ramp == "Field") 

data_controls$Time.point <- factor(data_controls$Time.point,
                       levels = c('Field','Stress', 'Recovery'),ordered = TRUE)

ggplot(data=data_controls, aes(x=Site, y = AverageRed, fill = Time.point))+ 
  geom_boxplot() + theme_minimal() + scale_fill_manual(values = c("grey55", "sienna1", "coral4")) + 
  ylab("Average Red Intensity \n (Higher = more bleached)") + xlab("") + theme(axis.text=element_text(size=14))
#blue shows stress, red shows recovery
#each point is one coral 
#lower value = less bleached 
#cannery and fagaalu controls did not bleach but tele and vatia did 

#ggsave("../../CSUMB/Thesis/Data analysis/Figures - bleaching phys/10232020/AvgRed_controls.pdf", height = 5, width = 8)

data_hightemp <- data %>%
  filter(data$Ramp == "35") 

data_medtemp <- data %>%
  filter(data$Ramp == "34") 

data_lowtemp <- data %>%
  filter(data$Ramp == "33") 

data_hightemp$Time.point <- factor(data_hightemp$Time.point, levels = c('Stress', 'Recovery'),ordered = TRUE)
data_medtemp$Time.point <- factor(data_hightemp$Time.point, levels = c('Stress', 'Recovery'),ordered = TRUE)
data_lowtemp$Time.point <- factor(data_hightemp$Time.point, levels = c('Stress', 'Recovery'),ordered = TRUE)

ggplot(data=data_hightemp, aes(x=Site, y = AverageRed, fill = Time.point))+ 
  geom_boxplot() + theme_minimal() + scale_fill_manual(values = c("sienna1", "coral4")) + 
  ylab("Average Red Intensity at 35C \n (Higher = more bleached)") + xlab("") + theme(axis.text=element_text(size=14))

#ggsave("../../CSUMB/Thesis/Data analysis/Figures - bleaching phys/10232020/AvgRed_hightemps.pdf", height = 5, width = 8)

ggplot(data=data_medtemp, aes(x=Site, y = AverageRed, fill = Time.point))+ 
  geom_boxplot() + theme_minimal() + scale_fill_manual(values = c("sienna1", "coral4")) + 
  ylab("Average Red Intensity at 34C \n (Higher = more bleached)") + xlab("") + theme(axis.text=element_text(size=14))

#ggsave("../../CSUMB/Thesis/Data analysis/Figures - bleaching phys/10232020/AvgRed_medtemps.pdf", height = 5, width = 8)

ggplot(data=data_lowtemp, aes(x=Site, y = AverageRed, fill = Time.point))+ 
  geom_boxplot() + theme_minimal() + scale_fill_manual(values = c("sienna1", "coral4")) + 
  ylab("Average Red Intensity at 33C \n (Higher = more bleached)") + xlab("") + theme(axis.text=element_text(size=14))

#ggsave("../../CSUMB/Thesis/Data analysis/Figures - bleaching phys/10232020/AvgRed_lowtemps.pdf", height = 5, width = 8)
```

Normalized colorimetric data to tank control at 28C for each colony 

Makes 3 plots:
1. 33C normalized to controls
2. 34C normalized to controls
3. 35C normalized to controls

```{r avgreddata_normalized}
head(data)
data$Coral.Watch.Color1 <- NULL
data$Coral.Watch.Color2 <- NULL
data$Coral.Watch.Color3 <- NULL
data$PAM..Fv.Fm. <- NULL
data$X <- NULL
#if tank A-D, rep 1; if tank E-F, rep 2
data$Rep[data$Tank == "A"] <- "1"
data$Rep[data$Tank == "B"] <- "1"
data$Rep[data$Tank == "C"] <- "1"
data$Rep[data$Tank == "D"] <- "1"
data$Rep[data$Tank == "E"] <- "2"
data$Rep[data$Tank == "F"] <- "2"
data$Rep[data$Tank == "G"] <- "2"
data$Rep[data$Tank == "H"] <- "2"
data$Tank <- NULL
data_normalized <- spread(data, Ramp, AverageRed)
head(data_normalized)
data_normalized$t35_normalizedto28 <- data_normalized$`35` - data_normalized$`28`
data_normalized$t34_normalizedto28 <- data_normalized$`34` - data_normalized$`28`
data_normalized$t33_normalizedto28 <- data_normalized$`33` - data_normalized$`28`
head(data_normalized)

data_normalized$Time.point <- factor(data_normalized$Time.point,
                                   levels = c('Stress', 'Recovery'),ordered = TRUE)

ggplot(data=data_normalized, aes(x=Site, y = t33_normalizedto28, fill = Time.point)) + geom_boxplot() + theme_minimal() + scale_fill_manual(values = c("sienna1", "coral4")) + 
  ylab("Average Red Intensity \n 33C normalized to 28C \n (Higher = more bleached)") + xlab("") + theme(axis.text=element_text(size=14))

#ggsave("../../CSUMB/Thesis/Data analysis/Figures - bleaching phys/10232020/AvgRed_lowtemps_normalized.pdf", height = 5, width = 8)

ggplot(data=data_normalized, aes(x=Site, y = t34_normalizedto28, fill = Time.point))+ geom_boxplot() + theme_minimal() + scale_fill_manual(values = c("sienna1", "coral4")) + 
  ylab("Average Red Intensity \n 34C normalized to 28C \n (Higher = more bleached)") + xlab("") + theme(axis.text=element_text(size=14))

#ggsave("../../CSUMB/Thesis/Data analysis/Figures - bleaching phys/10232020/AvgRed_medtemps_normalized.pdf", height = 5, width = 8)

ggplot(data=data_normalized, aes(x=Site, y = t35_normalizedto28, fill = Time.point))+ geom_boxplot() + theme_minimal() + scale_fill_manual(values = c("sienna1", "coral4")) + 
  ylab("Average Red Intensity \n 35C normalized to 28C \n (Higher = more bleached)") + xlab("") + theme(axis.text=element_text(size=14))

#ggsave("../../CSUMB/Thesis/Data analysis/Figures - bleaching phys/10232020/AvgRed_hightemps_normalized.pdf", height = 5, width = 8)
```

# CoralWatch Color Card



Color Card data - raw

Makes  figs
1. Control (28C) color score at time point 1
2. Control (28C) color score at time point 2
3. Control (28C) color score at time point 3
4. High (35C) color score at time point 1
5. High (35C) color score at time point 2
6. High (35C) color score at time point 3

```{r colorcard}
rm( list = ls())
graphics.off()
data <- read.csv("raw data sheets/MasterASData.csv")
head(data)
data$Time.point <- as.character(data$Time.point)
data$Time.point[data$Ramp == "Field"] <- "Field"

data$CWchange1 <- data$Coral.Watch.Color1 - data$Coral.Watch.Color2
data$CWchange2 <- data$Coral.Watch.Color1 - data$Coral.Watch.Color3

data$Time.point <- factor(data$Time.point, levels = c('Field', 'Stress', 'Recovery'),ordered = TRUE)

data_controls <- data %>%
  filter(data$Ramp == "28" | data$Ramp == "Field") 

#raw data of controls
ggplot(data=data_controls, aes(x=Site, y = Coral.Watch.Color1, fill = Time.point))+ 
  geom_boxplot() + theme_minimal() + scale_fill_manual(values = c("grey55", "sienna1", "coral4")) + 
  ylab("Color Score of controls at T=1") + xlab("") + theme(axis.text=element_text(size=14))

ggplot(data=data_controls, aes(x=Site, y = Coral.Watch.Color2, fill = Time.point))+ 
  geom_boxplot() + theme_minimal() + scale_fill_manual(values = c("sienna1", "coral4")) + 
  ylab("Color Score of controls at T=2") + xlab("") + theme(axis.text=element_text(size=14))

ggplot(data=data_controls, aes(x=Site, y = Coral.Watch.Color3, fill = Time.point))+ 
  geom_boxplot() + theme_minimal() + scale_fill_manual(values = c("coral4")) + 
  ylab("Color Score of controls at T=3") + xlab("") + theme(axis.text=element_text(size=14))

#raw data of high temp 35C

data_hightemp <- data %>%
  filter(data$Ramp == "35") 

ggplot(data=data_hightemp, aes(x=Site, y = Coral.Watch.Color1, fill = Time.point))+ 
  geom_boxplot() + theme_minimal() + scale_fill_manual(values = c("sienna1", "coral4")) + 
  ylab("Color Score of 35C at T=1") + xlab("") + theme(axis.text=element_text(size=14))


ggplot(data=data_hightemp, aes(x=Site, y = Coral.Watch.Color2, fill = Time.point))+ 
  geom_boxplot() + theme_minimal() + scale_fill_manual(values = c("sienna1", "coral4")) + 
  ylab("Color Score of 35C at T=2") + xlab("") + theme(axis.text=element_text(size=14))

ggplot(data=data_hightemp, aes(x=Site, y = Coral.Watch.Color3, fill = Time.point))+ 
  geom_boxplot() + theme_minimal() + scale_fill_manual(values = c("coral4")) + 
  ylab("Color Score of 35C at T=3") + xlab("") + theme(axis.text=element_text(size=14))

#change in color over time

#from Cw2 to CW1
ggplot(data=data_hightemp, aes(x=Site, y = CWchange1, fill = Site))+ 
  geom_boxplot() + theme_minimal()  + 
  ylab("Color Score of 35C at T1 - T2 \n (higher = more bleached, lower = less bleached)") + xlab("") + theme(axis.text=element_text(size=14))

#ggsave("../../CSUMB/Thesis/Data analysis/Figures - bleaching phys/10232020/CW_T1-T2_raw.pdf", height = 5, width = 8)

#from CW3 to CW1
ggplot(data=data_hightemp, aes(x=Site, y = CWchange2, fill = Site))+ 
  geom_boxplot() + theme_minimal()  + 
  ylab("Color Score of 35C at T1 - T3 \n (higher = more bleached, lower = less bleached)") + xlab("") + theme(axis.text=element_text(size=14))

#ggsave("../../CSUMB/Thesis/Data analysis/Figures - bleaching phys/10232020/CW_T1-T3_raw.pdf", height = 5, width = 8)

```

Colorimetric data normalized to controls (at 28C)



```{r colorimetricdata_normalized}

data <- read.csv("raw data sheets/MasterASData.csv")
data_cw <- data
head(data_cw)
data_cw$AverageRed <- NULL
data_cw$PAM..Fv.Fm. <- NULL
data_cw$X <- NULL
#if tank A-D, rep 1; if tank E-F, rep 2
data_cw$Rep[data_cw$Tank == "A"] <- "1"
data_cw$Rep[data_cw$Tank == "B"] <- "1"
data_cw$Rep[data_cw$Tank == "C"] <- "1"
data_cw$Rep[data_cw$Tank == "D"] <- "1"
data_cw$Rep[data_cw$Tank == "E"] <- "2"
data_cw$Rep[data_cw$Tank == "F"] <- "2"
data_cw$Rep[data_cw$Tank == "G"] <- "2"
data_cw$Rep[data_cw$Tank == "H"] <- "2"
data_cw$Tank <- NULL
data_cw
head(data_cw)
data_cw_1 <- data_cw
data_cw_2 <- data_cw
data_cw_3 <- data_cw
data_cw_1$Coral.Watch.Color2 <- NULL
data_cw_1$Coral.Watch.Color3 <- NULL
data_cw_2$Coral.Watch.Color1 <- NULL
data_cw_2$Coral.Watch.Color3 <- NULL
data_cw_3$Coral.Watch.Color1 <- NULL
data_cw_3$Coral.Watch.Color2 <- NULL

data_normalized_CW1 <- spread(data_cw_1, Ramp, Coral.Watch.Color1)
head(data_normalized_CW1)
data_normalized_CW2 <- spread(data_cw_2, Ramp, Coral.Watch.Color2)
head(data_normalized_CW2)
data_normalized_CW3 <- spread(data_cw_3, Ramp, Coral.Watch.Color3)
head(data_normalized_CW3)


data_normalized_CW2$t35_normalizedto28 <- data_normalized_CW2$`28` - data_normalized_CW2$`35`
data_normalized_CW3$t35_normalizedto28 <- data_normalized_CW3$`28` - data_normalized_CW3$`35`

ggplot(data=data_normalized_CW2, aes(x=Site, y = t35_normalizedto28, fill = Site))+ 
  geom_boxplot() + theme_minimal() + 
  ylab("Color Score of 35C Normalized to Control at T=2 \n (higher = more bleached)") + xlab("") + theme(axis.text=element_text(size=14))

ggsave("../../CSUMB/Thesis/Data analysis/Figures - bleaching phys/10232020/CW_T2_35C_normalized.pdf", height = 5, width = 8)


ggplot(data=data_normalized_CW3, aes(x=Site, y = t35_normalizedto28, fill = Site))+ 
  geom_boxplot() + theme_minimal() + 
  ylab("Color Score of 35C Normalized to Control at T=3 \n (higher = more bleached)") + xlab("") + theme(axis.text=element_text(size=14))

ggsave("../../CSUMB/Thesis/Data analysis/Figures - bleaching phys/10232020/CW_T3_35C_normalized.pdf", height = 5, width = 8)


```


#PAM Fv/Fm



PAM data

Raw controls and 35 data 

```{r PAM}
rm( list = ls())
graphics.off()
data <- read.csv("raw data sheets/MasterASData.csv")
head(data)
data$Time.point <- as.character(data$Time.point)
data$Time.point[data$Ramp == "Field"] <- "Field"


##########work with this one for PAM
ggplot(data=data, aes(x=Site, y = PAM..Fv.Fm., fill = Ramp))+ 
  geom_boxplot() + theme_minimal() + 
  ylab("PAM Fv/Fm (higher = healthier symbionts)") + xlab("") + theme(axis.text=element_text(size=14))+ ylim(0.5,0.7)


data_controls <- data %>%
  filter(data$Ramp == "28" | data$Ramp == "Field") 

ggplot(data=data_controls, aes(x=Site, y = PAM..Fv.Fm., fill = Site))+ 
  geom_boxplot() + theme_minimal() + scale_fill_manual(values = c("grey55", "sienna1", "coral4")) + 
  ylab("PAM Fv/Fm (higher = healthier symbionts)") + xlab("") + theme(axis.text=element_text(size=14))

data_hightemp <- data %>%
  filter(data$Ramp == "35") 

ggplot(data=data_hightemp, aes(x=Site, y = PAM..Fv.Fm., fill = Site))+ 
  geom_boxplot() + theme_minimal() + scale_fill_manual(values = c("sienna1", "coral4")) + ylab("PAM Fv/Fm (higher = healthier symbionts)") + xlab("") + theme(axis.text=element_text(size=14)) + ylim(0.5,0.7)


```




```{r PAM_normalized}
rm( list = ls())
graphics.off()
data <- read.csv("raw data sheets/MasterASData.csv")
head(data)
data$Time.point <- as.character(data$Time.point)
data$Time.point[data$Ramp == "Field"] <- "Field"


data$Coral.Watch.Color1 <- NULL
data$Coral.Watch.Color2 <- NULL
data$Coral.Watch.Color3 <- NULL
data$AverageRed <- NULL
data$X <- NULL
#if tank A-D, rep 1; if tank E-F, rep 2
data$Rep[data$Tank == "A"] <- "1"
data$Rep[data$Tank == "B"] <- "1"
data$Rep[data$Tank == "C"] <- "1"
data$Rep[data$Tank == "D"] <- "1"
data$Rep[data$Tank == "E"] <- "2"
data$Rep[data$Tank == "F"] <- "2"
data$Rep[data$Tank == "G"] <- "2"
data$Rep[data$Tank == "H"] <- "2"
data$Tank <- NULL
data_normalized <- spread(data, Ramp, PAM..Fv.Fm.)
head(data_normalized)
data_normalized$t35_normalizedto28 <- data_normalized$`35` - data_normalized$`28`
data_normalized$t34_normalizedto28 <- data_normalized$`34` - data_normalized$`28`
data_normalized$t33_normalizedto28 <- data_normalized$`33` - data_normalized$`28`
head(data_normalized)

#plots not showing outliers 

ggplot(data=data_normalized, aes(x=Site, y = t33_normalizedto28, fill = Site)) + geom_boxplot() + theme_minimal() + 
  ylab("PAM Fv/Fm \n 33C normalized to 28C \n (higher = healthier symbionts)") + xlab("") + theme(axis.text=element_text(size=14)) + ylim(-0.15,0.1)

ggsave("../../CSUMB/Thesis/Data analysis/Figures - bleaching phys/10232020/PAM_lowtemps_normalized.pdf", height = 5, width = 8)

ggplot(data=data_normalized, aes(x=Site, y = t34_normalizedto28, fill = Site))+ geom_boxplot() + theme_minimal() + 
  ylab("PAM Fv/Fm \n 34C normalized to 28C \n (higher = healthier symbionts)") + xlab("") + theme(axis.text=element_text(size=14)) + ylim(-0.15,0.15)

ggsave("../../CSUMB/Thesis/Data analysis/Figures - bleaching phys/10232020/PAM_medtemps_normalized.pdf", height = 5, width = 8)

ggplot(data=data_normalized, aes(x=Site, y = t35_normalizedto28, fill = Site))+ geom_boxplot() + theme_minimal()  + 
  ylab("PAM Fv/Fm \n 35C normalized to 28C \n (higher = healthier symbionts)") + xlab("") + theme(axis.text=element_text(size=14)) + ylim(-0.15,0.15)

ggsave("../../CSUMB/Thesis/Data analysis/Figures - bleaching phys/10232020/PAM_hightemps_normalized.pdf", height = 5, width = 8)


```

```{r PAM_stats}
library(car)
library(lmerTest)
library(emmeans)
rm( list = ls())
graphics.off()
data <- read.csv("raw data sheets/MasterASData.csv")
head(data)
data$Time.point <- as.character(data$Time.point)
data$Time.point[data$Ramp == "Field"] <- "Field"
data$PAM..Fv.Fm.[data$PAM..Fv.Fm. == 0] <- NA #remove dead corals (0 Fv/Fm)
head(data)

# Mixed Effects model - random effect of replicate.group
pam.cbass.model <- lmer(PAM..Fv.Fm. ~ Site * Ramp + (1|Tank), data = data)
anova(pam.cbass.model) # anova from lmerTest ANOVA SS Type III with ddf="Satterthwaite"
ranova(pam.cbass.model) # anova-like table for random effects

ggplot(data=data, aes(x=Site, y = PAM..Fv.Fm., fill = Tank))+ geom_boxplot() + theme_minimal()  

# Model fitting and assumptions diagnostic 
plot(PAM..Fv.Fm. ~ interaction(Site,Ramp,Tank), data = data) 
# Box-plot homogeinity of variance
leveneTest(PAM..Fv.Fm. ~ Site * Ramp * Tank, data=data)
# formal statistical test for homogeinity of variance (not recommended due the small sample size)
#not signif! 
plot(pam.cbass.model) # Residual vs Fitted values 
#looks fine I think
qqnorm(resid(pam.cbass.model)); qqline(resid(pam.cbass.model)) 
# qq plot to check for normal distribution of residuals
#mostly okay?
hist(resid(pam.cbass.model)) 
# histogram of residuals to check for normal distribution of residuals
shapiro.test(resid(pam.cbass.model)) 
# formal statistical test (not recommended due the small sample size)
#not normal at all

# comparing between reef sites within each temperature treatment
pam.cbass.emms.reef <- emmeans(pam.cbass.model, pairwise ~ Site|Ramp, weights = "proportional", adjust="none")
summary(pam.cbass.emms.reef$emmeans)

```



