---
title: "Physiology_figs"
author: "Melissa Naugle"
date: "10/23/2020"
output: html_document
---

How to use this document:
1. Navigate to whichever section you want to use (red intensity, color card, or PAM)
2. Begin with first chunk of that section. Each section can be run independently but not each chunk can be run independently (since data is read in in first chunk of each section)


```{r setup}
setwd(dir = "~/Desktop/GitHub/CBASS_bleachingdata/")
rm( list = ls())
graphics.off()
library(ggplot2)
require(plyr)
library(dplyr)
library(reshape2)
library(mosaic)
library(tidyr)
library(ggpubr)
```


# Red Intensity 


Plot raw colorimetric data

Makes 4 plots:
1. controls (tank (28C) and field controls)
2. 35C
3. 34C
4. 33C

```{r avgreddata}
data <- read.csv("raw data sheets/MasterASData.csv")
head(data)
data$Time.point <- as.character(data$Time.point)
data$Time.point[data$Ramp == "Field"] <- "Field"

#plot all temps at stress

data_stress <- data %>%
  filter(data$Time.point == "Stress") 

#ggplot(data=data_stress, aes(x=Site, y = AverageRed, fill = Ramp))+ geom_boxplot() + theme_minimal()  + scale_fill_manual(values = c("dodgerblue4", "cyan4", "springgreen3", "yellow2"))  + ylab("Average Red Intensity \n Stress Time Point \n (Higher = more bleached)") + xlab("") + theme(axis.text=element_text(size=14))

#ggsave("../../CSUMB/Thesis/Data analysis/Figures - bleaching phys/10232020/AvgRed_stress.pdf", height = 5, width = 8)

#plot all temps at recovery 

data_recovery <- data %>%
  filter(data$Time.point == "Recovery") 

#ggplot(data=data_recovery, aes(x=Site, y = AverageRed, fill = Ramp))+ geom_boxplot() + theme_minimal()  + scale_fill_manual(values = c("dodgerblue4", "cyan4", "springgreen3", "yellow2"))  +ylab("Average Red Intensity \n Recovery Time Point \n (Higher = more bleached)") + xlab("") + theme(axis.text=element_text(size=14))

#ggsave("../../CSUMB/Thesis/Data analysis/Figures - bleaching phys/10232020/AvgRed_recovery.pdf", height = 5, width = 8)




#filter out only controls
data_controls <- data %>%
  filter(data$Ramp == "28" | data$Ramp == "Field") 

data_controls$Time.point <- factor(data_controls$Time.point,
                       levels = c('Field','Stress', 'Recovery'),ordered = TRUE)

#ggplot(data=data_controls, aes(x=Site, y = AverageRed, fill = Time.point))+ geom_boxplot() + theme_minimal() + scale_fill_manual(values = c("grey55", "sienna1", "coral4")) + ylab("Average Red Intensity \n (Higher = more bleached)") + xlab("") + theme(axis.text=element_text(size=14))

#blue shows stress, red shows recovery
#each point is one coral 
#lower value = less bleached 
#cannery and fagaalu controls did not bleach but tele and vatia did 

#ggsave("../../CSUMB/Thesis/Data analysis/Figures - bleaching phys/10232020/AvgRed_controls.pdf", height = 5, width = 8)

data_hightemp <- data %>%
  filter(data$Ramp == "35") 

data_medtemp <- data %>%
  filter(data$Ramp == "34") 

data_lowtemp <- data %>%
  filter(data$Ramp == "33") 

data_hightemp$Time.point <- factor(data_hightemp$Time.point, levels = c('Stress', 'Recovery'),ordered = TRUE)
data_medtemp$Time.point <- factor(data_hightemp$Time.point, levels = c('Stress', 'Recovery'),ordered = TRUE)
data_lowtemp$Time.point <- factor(data_hightemp$Time.point, levels = c('Stress', 'Recovery'),ordered = TRUE)

#ggplot(data=data_hightemp, aes(x=Site, y = AverageRed, fill = Time.point))+  geom_boxplot() + theme_minimal() + scale_fill_manual(values = c("sienna1", "coral4")) + ylab("Average Red Intensity at 35C \n (Higher = more bleached)") + xlab("") + theme(axis.text=element_text(size=14))

#ggsave("../../CSUMB/Thesis/Data analysis/Figures - bleaching phys/10232020/AvgRed_hightemps.pdf", height = 5, width = 8)

#ggplot(data=data_medtemp, aes(x=Site, y = AverageRed, fill = Time.point))+ geom_boxplot() + theme_minimal() + scale_fill_manual(values = c("sienna1", "coral4")) + ylab("Average Red Intensity at 34C \n (Higher = more bleached)") + xlab("") + theme(axis.text=element_text(size=14))

#ggsave("../../CSUMB/Thesis/Data analysis/Figures - bleaching phys/10232020/AvgRed_medtemps.pdf", height = 5, width = 8)

#ggplot(data=data_lowtemp, aes(x=Site, y = AverageRed, fill = Time.point))+ geom_boxplot() + theme_minimal() + scale_fill_manual(values = c("sienna1", "coral4")) + ylab("Average Red Intensity at 33C \n (Higher = more bleached)") + xlab("") + theme(axis.text=element_text(size=14))

#ggsave("../../CSUMB/Thesis/Data analysis/Figures - bleaching phys/10232020/AvgRed_lowtemps.pdf", height = 5, width = 8)
```

Normalized colorimetric data to tank control at 28C for each colony 

Makes 5 plots:
1. 33C normalized to controls
2. 34C normalized to controls
3. 35C normalized to controls
4. All temps normalized at stress
5. All temps normalized at recovery

```{r avgreddata_normalized}
head(data)
data$Coral.Watch.Color1 <- NULL
data$Coral.Watch.Color2 <- NULL
data$Coral.Watch.Color3 <- NULL
data$PAM..Fv.Fm. <- NULL
data$X <- NULL
#if tank A-D, rep 1; if tank E-F, rep 2
data$Rep[data$Tank == "A"] <- "1"
data$Rep[data$Tank == "B"] <- "1"
data$Rep[data$Tank == "C"] <- "1"
data$Rep[data$Tank == "D"] <- "1"
data$Rep[data$Tank == "E"] <- "2"
data$Rep[data$Tank == "F"] <- "2"
data$Rep[data$Tank == "G"] <- "2"
data$Rep[data$Tank == "H"] <- "2"
data$Tank <- NULL
data_normalized <- spread(data, Ramp, AverageRed)
head(data_normalized)
data_normalized$t35_normalizedto28 <- data_normalized$`35` - data_normalized$`28`
data_normalized$t34_normalizedto28 <- data_normalized$`34` - data_normalized$`28`
data_normalized$t33_normalizedto28 <- data_normalized$`33` - data_normalized$`28`
head(data_normalized)

data_normalized$Time.point <- factor(data_normalized$Time.point,
                                   levels = c('Stress', 'Recovery'),ordered = TRUE)

#ggplot(data=data_normalized, aes(x=Site, y = t33_normalizedto28, fill = Time.point)) + geom_boxplot() + theme_minimal() + scale_fill_manual(values = c("sienna1", "coral4")) + ylab("Average Red Intensity \n 33C normalized to 28C \n (Higher = more bleached)") + xlab("") + theme(axis.text=element_text(size=14))

#ggsave("../../CSUMB/Thesis/Data analysis/Figures - bleaching phys/10232020/AvgRed_lowtemps_normalized.pdf", height = 5, width = 8)

#ggplot(data=data_normalized, aes(x=Site, y = t34_normalizedto28, fill = Time.point))+ geom_boxplot() + theme_minimal() + scale_fill_manual(values = c("sienna1", "coral4")) + ylab("Average Red Intensity \n 34C normalized to 28C \n (Higher = more bleached)") + xlab("") + theme(axis.text=element_text(size=14))

#ggsave("../../CSUMB/Thesis/Data analysis/Figures - bleaching phys/10232020/AvgRed_medtemps_normalized.pdf", height = 5, width = 8)

#ggplot(data=data_normalized, aes(x=Site, y = t35_normalizedto28, fill = Time.point))+ geom_boxplot() + theme_minimal() + scale_fill_manual(values = c("sienna1", "coral4")) + ylab("Average Red Intensity \n 35C normalized to 28C \n (Higher = more bleached)") + xlab("") + theme(axis.text=element_text(size=14))

#ggsave("../../CSUMB/Thesis/Data analysis/Figures - bleaching phys/10232020/AvgRed_hightemps_normalized.pdf", height = 5, width = 8)

#all in one plot 

data_normalized$StressRecovTimePoint <- data_normalized$Time.point
data_normalized$Time.point <- NULL

data_normalized_gather <- pivot_longer(data_normalized, cols = starts_with("t"), names_to = "temperature_norm", values_to = "averagered" )

data_normalized_gather_stress <- data_normalized_gather[data_normalized_gather$StressRecovTimePoint == "Stress",]
data_normalized_gather_recovery <- data_normalized_gather[data_normalized_gather$StressRecovTimePoint == "Recovery",]


data_normalized_gather_stress_nas <- data_normalized_gather_stress[!is.na(data_normalized_gather_stress$averagered),]
  
data_normalized_gather_recovery_nas <- data_normalized_gather_recovery[!is.na(data_normalized_gather_recovery$averagered),]
 
  
#ggplot(data=data_normalized_gather_stress_nas, aes(x=Site, y = averagered, fill = temperature_norm))+ geom_boxplot() + theme_minimal() + scale_fill_manual(name = "Temperature", labels = c("33C", "34C", "35C"), values = c("mediumseagreen", "darkcyan", "darkgoldenrod")) + ylab("Average Red Intensity \n (normalized to 28C)") + xlab("") + theme(axis.text=element_text(size=14)) 

#ggsave("../../CSUMB/Thesis/Data analysis/Figures - bleaching phys/02152021- normalized to controls/AvgRed_stress_normalized.pdf", height = 5, width = 8)

#ggplot(data=data_normalized_gather_recovery_nas, aes(x=Site, y = averagered, fill = temperature_norm))+ geom_boxplot() + theme_minimal() + scale_fill_manual(name = "Temperature", labels = c("33C", "34C", "35C"), values = c("mediumseagreen", "darkcyan", "darkgoldenrod")) + ylab("Average Red Intensity \n (normalized to controls)") + xlab("") + theme(axis.text=element_text(size=14))

#ggsave("../../CSUMB/Thesis/Data analysis/Figures - bleaching phys/02152021- normalized to controls/AvgRed_recov_normalized.pdf", height = 5, width = 8)

red_stress <- ggplot(data=data_normalized_gather_stress_nas, aes(x=Site, y = averagered, fill = temperature_norm))+ geom_boxplot(outlier.size = -1) + theme_minimal() + scale_fill_manual(name = "Temperature", labels = c("33C", "34C", "35C"), values = c("mediumseagreen", "darkcyan", "darkgoldenrod")) + ylab("Average Red Intensity") + xlab("") + theme(axis.text=element_text(size=14)) + ggtitle("Heat Stress") + theme(plot.title = element_text(hjust = 0.5))

red_recov <- ggplot(data=data_normalized_gather_recovery_nas, aes(x=Site, y = averagered, fill = temperature_norm))+ geom_boxplot(outlier.size = -1) + theme_minimal() + scale_fill_manual(name = "Temperature", labels = c("33C", "34C", "35C"), values = c("mediumseagreen", "darkcyan", "darkgoldenrod")) + ylab("Average Red Intensity") + xlab("") + theme(axis.text=element_text(size=14)) + ggtitle("Recovery") + theme(plot.title = element_text(hjust = 0.5))

ggarrange(red_stress, red_recov, ncol = 1, nrow = 2, common.legend = TRUE, legend = "right")

#ggsave("../../CSUMB/Thesis/Data analysis/Figures - bleaching phys/02152021- normalized to controls/Avgred_normalized_stressrecov.pdf", height = 9, width = 8)
```

# CoralWatch Color Card



Color Card data - raw

Makes  figs
1. Control (28C) color score at time point 1
2. Control (28C) color score at time point 2
3. Control (28C) color score at time point 3
4. High (35C) color score at time point 1
5. High (35C) color score at time point 2
6. High (35C) color score at time point 3

```{r colorcard}
rm( list = ls())
graphics.off()
data <- read.csv("raw data sheets/MasterASData.csv")
head(data)
data$Time.point <- as.character(data$Time.point)
data$Time.point[data$Ramp == "Field"] <- "Field"

data$CWchange1 <- data$Coral.Watch.Color1 - data$Coral.Watch.Color2
data$CWchange2 <- data$Coral.Watch.Color1 - data$Coral.Watch.Color3

data$Time.point <- factor(data$Time.point, levels = c('Field', 'Stress', 'Recovery'),ordered = TRUE)

ggplot(data=data, aes(x=Site, y = CWchange1, fill = Ramp))+ 
  geom_boxplot() + theme_minimal() + scale_fill_manual(values = c("dodgerblue4", "cyan4", "springgreen3", "yellow2")) + 
  ylab("Color Score T1 - T2 \n (higher = more bleached, lower = less bleached)") + xlab("") + theme(axis.text=element_text(size=14))

ggsave("../../CSUMB/Thesis/Data analysis/Figures - bleaching phys/10232020/CW_all_T1-T2_raw.pdf", height = 5, width = 8)

ggplot(data=data, aes(x=Site, y = CWchange2, fill = Ramp))+ 
  geom_boxplot() + theme_minimal() + scale_fill_manual(values = c("dodgerblue4", "cyan4", "springgreen3", "yellow2")) + 
  ylab("Color Score T1 - T3 \n (higher = more bleached, lower = less bleached)") + xlab("") + theme(axis.text=element_text(size=14))

ggsave("../../CSUMB/Thesis/Data analysis/Figures - bleaching phys/10232020/CW_all_T1-T3_raw.pdf", height = 5, width = 8)


data_controls <- data %>%
  filter(data$Ramp == "28" | data$Ramp == "Field") 

#raw data of controls
ggplot(data=data_controls, aes(x=Site, y = Coral.Watch.Color1, fill = Time.point))+ 
  geom_boxplot() + theme_minimal() + scale_fill_manual(values = c("grey55", "sienna1", "coral4")) + 
  ylab("Color Score of controls at T=1") + xlab("") + theme(axis.text=element_text(size=14))

ggplot(data=data_controls, aes(x=Site, y = Coral.Watch.Color2, fill = Time.point))+ 
  geom_boxplot() + theme_minimal() + scale_fill_manual(values = c("sienna1", "coral4")) + 
  ylab("Color Score of controls at T=2") + xlab("") + theme(axis.text=element_text(size=14))

ggplot(data=data_controls, aes(x=Site, y = Coral.Watch.Color3, fill = Time.point))+ 
  geom_boxplot() + theme_minimal() + scale_fill_manual(values = c("coral4")) + 
  ylab("Color Score of controls at T=3") + xlab("") + theme(axis.text=element_text(size=14))

#raw data of high temp 35C

data_hightemp <- data %>%
  filter(data$Ramp == "35") 

ggplot(data=data_hightemp, aes(x=Site, y = Coral.Watch.Color1, fill = Time.point))+ 
  geom_boxplot() + theme_minimal() + scale_fill_manual(values = c("sienna1", "coral4")) + 
  ylab("Color Score of 35C at T=1") + xlab("") + theme(axis.text=element_text(size=14))


ggplot(data=data_hightemp, aes(x=Site, y = Coral.Watch.Color2, fill = Time.point))+ 
  geom_boxplot() + theme_minimal() + scale_fill_manual(values = c("sienna1", "coral4")) + 
  ylab("Color Score of 35C at T=2") + xlab("") + theme(axis.text=element_text(size=14))

ggplot(data=data_hightemp, aes(x=Site, y = Coral.Watch.Color3, fill = Time.point))+ 
  geom_boxplot() + theme_minimal() + scale_fill_manual(values = c("coral4")) + 
  ylab("Color Score of 35C at T=3") + xlab("") + theme(axis.text=element_text(size=14))

#change in color over time

#from Cw2 to CW1
ggplot(data=data_hightemp, aes(x=Site, y = CWchange1, fill = Site))+ 
  geom_boxplot() + theme_minimal()  + 
  ylab("Color Score of 35C at T1 - T2 \n (higher = more bleached, lower = less bleached)") + xlab("") + theme(axis.text=element_text(size=14))

#ggsave("../../CSUMB/Thesis/Data analysis/Figures - bleaching phys/10232020/CW_T1-T2_raw.pdf", height = 5, width = 8)

#from CW3 to CW1
ggplot(data=data_hightemp, aes(x=Site, y = CWchange2, fill = Site))+ 
  geom_boxplot() + theme_minimal()  + 
  ylab("Color Score of 35C at T1 - T3 \n (higher = more bleached, lower = less bleached)") + xlab("") + theme(axis.text=element_text(size=14))

#ggsave("../../CSUMB/Thesis/Data analysis/Figures - bleaching phys/10232020/CW_T1-T3_raw.pdf", height = 5, width = 8)

```

Colorimetric data normalized to controls (at 28C)



```{r colorimetricdata_normalized}

data <- read.csv("raw data sheets/MasterASData.csv")
data_cw <- data
head(data_cw)
data_cw$AverageRed <- NULL
data_cw$PAM..Fv.Fm. <- NULL
data_cw$X <- NULL
#if tank A-D, rep 1; if tank E-F, rep 2
data_cw$Rep[data_cw$Tank == "A"] <- "1"
data_cw$Rep[data_cw$Tank == "B"] <- "1"
data_cw$Rep[data_cw$Tank == "C"] <- "1"
data_cw$Rep[data_cw$Tank == "D"] <- "1"
data_cw$Rep[data_cw$Tank == "E"] <- "2"
data_cw$Rep[data_cw$Tank == "F"] <- "2"
data_cw$Rep[data_cw$Tank == "G"] <- "2"
data_cw$Rep[data_cw$Tank == "H"] <- "2"
data_cw$Tank <- NULL
data_cw
head(data_cw)
data_cw_1 <- data_cw
data_cw_2 <- data_cw
data_cw_3 <- data_cw
data_cw_1$Coral.Watch.Color2 <- NULL
data_cw_1$Coral.Watch.Color3 <- NULL
data_cw_2$Coral.Watch.Color1 <- NULL
data_cw_2$Coral.Watch.Color3 <- NULL
data_cw_3$Coral.Watch.Color1 <- NULL
data_cw_3$Coral.Watch.Color2 <- NULL

data_normalized_CW1 <- spread(data_cw_1, Ramp, Coral.Watch.Color1)
head(data_normalized_CW1)
data_normalized_CW2 <- spread(data_cw_2, Ramp, Coral.Watch.Color2)
head(data_normalized_CW2)
data_normalized_CW3 <- spread(data_cw_3, Ramp, Coral.Watch.Color3)
head(data_normalized_CW3)


data_normalized_CW2$t35_normalizedto28 <- data_normalized_CW2$`35` - data_normalized_CW2$`28`
data_normalized_CW2$t34_normalizedto28 <- data_normalized_CW2$`34` - data_normalized_CW2$`28`
data_normalized_CW2$t33_normalizedto28 <- data_normalized_CW2$`33` - data_normalized_CW2$`28`
data_normalized_CW3$t35_normalizedto28 <- data_normalized_CW3$`35` - data_normalized_CW3$`28`
data_normalized_CW3$t34_normalizedto28 <- data_normalized_CW3$`34` - data_normalized_CW3$`28`
data_normalized_CW3$t33_normalizedto28 <- data_normalized_CW3$`33` - data_normalized_CW3$`28`



#ggplot(data=data_normalized_CW2, aes(x=Site, y = t35_normalizedto28, fill = Site))+ geom_boxplot() + theme_minimal() +ylab("Color Score of 35C Normalized to Control at T=2 \n (higher = more bleached)") + xlab("") + theme(axis.text=element_text(size=14))

#ggsave("../../CSUMB/Thesis/Data analysis/Figures - bleaching phys/10232020/CW_T2_35C_normalized.pdf", height = 5, width = 8)


#ggplot(data=data_normalized_CW3, aes(x=Site, y = t35_normalizedto28, fill = Site))+ geom_boxplot() + theme_minimal() +  ylab("Color Score of 35C Normalized to Control at T=3 \n (higher = more bleached)") + xlab("") + theme(axis.text=element_text(size=14))

#ggsave("../../CSUMB/Thesis/Data analysis/Figures - bleaching phys/10232020/CW_T3_35C_normalized.pdf", height = 5, width = 8)



#all in one plot

data_normalized <- data_normalized_CW3

data_normalized$StressRecovTimePoint <- data_normalized$Time.point
data_normalized$Time.point <- NULL

data_normalized_gather <- pivot_longer(data_normalized, cols = starts_with("t"), names_to = "temperature_norm", values_to = "colorcard" )

data_normalized_gather_nas <- data_normalized_gather[!is.na(data_normalized_gather$colorcard),]
  
  
ggplot(data=data_normalized_gather_nas, aes(x=Site, y = colorcard, fill = temperature_norm))+ geom_boxplot(outlier.size = -1) + theme_minimal() + scale_fill_manual(name = "Temperature", labels = c("33C", "34C", "35C"), values = c("mediumseagreen", "darkcyan", "darkgoldenrod")) + ylab("Color Card Score") + xlab("") + theme(axis.text=element_text(size=14)) 

#ggsave("../../CSUMB/Thesis/Data analysis/Figures - bleaching phys/02152021- normalized to controls/colorcard_t3_normalized.pdf", height = 5, width = 8)

data_normalized <- data_normalized_CW2

data_normalized$StressRecovTimePoint <- data_normalized$Time.point
data_normalized$Time.point <- NULL

data_normalized_gather <- pivot_longer(data_normalized, cols = starts_with("t"), names_to = "temperature_norm", values_to = "colorcard" )

data_normalized_gather_nas <- data_normalized_gather[!is.na(data_normalized_gather$colorcard),]

ggplot(data=data_normalized_gather_nas, aes(x=Site, y = colorcard, fill = temperature_norm))+ geom_boxplot(outlier.size = -1) + theme_minimal() + scale_fill_manual(name = "Temperature", labels = c("33C", "34C", "35C"), values = c("mediumseagreen", "darkcyan", "darkgoldenrod")) + ylab("Color Card Score") + xlab("") + theme(axis.text=element_text(size=14)) 

#ggsave("../../CSUMB/Thesis/Data analysis/Figures - bleaching phys/02152021- normalized to controls/colorcard_t2_normalized.pdf", height = 5, width = 8)

```


#PAM Fv/Fm



PAM data

Raw controls and 35 data 

```{r PAM}
rm( list = ls())
graphics.off()
data <- read.csv("raw data sheets/MasterASData.csv")
head(data)
data$Time.point <- as.character(data$Time.point)
data$Time.point[data$Ramp == "Field"] <- "Field"

data$Ramp <- factor(data$Ramp, levels = c('Field', '28', '33', '34', '35'),ordered = TRUE)

ggplot(data=data, aes(x=Site, y = PAM..Fv.Fm., fill = Ramp))+ 
  geom_boxplot() + theme_minimal() + 
  ylab("PAM Fv/Fm (higher = healthier symbionts)") + xlab("") + theme(axis.text=element_text(size=14))+ ylim(0.5,0.7) 


#ggsave("../../CSUMB/Thesis/Data analysis/Figures - bleaching phys/10232020/PAM_bysiteandtemp.pdf", height = 5, width = 8)

data_controls <- data %>%
  filter(data$Ramp == "28" | data$Ramp == "Field") 

ggplot(data=data_controls, aes(x=Site, y = PAM..Fv.Fm., fill = Site))+ 
  geom_boxplot() + theme_minimal() + scale_fill_manual(values = c("grey55", "sienna1", "coral4")) + 
  ylab("PAM Fv/Fm (higher = healthier symbionts)") + xlab("") + theme(axis.text=element_text(size=14))

data_hightemp <- data %>%
  filter(data$Ramp == "35") 

ggplot(data=data_hightemp, aes(x=Site, y = PAM..Fv.Fm., fill = Site))+ 
  geom_boxplot() + theme_minimal() + scale_fill_manual(values = c("sienna1", "coral4")) + ylab("PAM Fv/Fm (higher = healthier symbionts)") + xlab("") + theme(axis.text=element_text(size=14)) + ylim(0.5,0.7)


```




```{r PAM_normalized}
rm( list = ls())
graphics.off()
data <- read.csv("raw data sheets/MasterASData.csv")
head(data)
data$Time.point <- as.character(data$Time.point)
data$Time.point[data$Ramp == "Field"] <- "Field"


data$Coral.Watch.Color1 <- NULL
data$Coral.Watch.Color2 <- NULL
data$Coral.Watch.Color3 <- NULL
data$AverageRed <- NULL
data$X <- NULL
#if tank A-D, rep 1; if tank E-F, rep 2
data$Rep[data$Tank == "A"] <- "1"
data$Rep[data$Tank == "B"] <- "1"
data$Rep[data$Tank == "C"] <- "1"
data$Rep[data$Tank == "D"] <- "1"
data$Rep[data$Tank == "E"] <- "2"
data$Rep[data$Tank == "F"] <- "2"
data$Rep[data$Tank == "G"] <- "2"
data$Rep[data$Tank == "H"] <- "2"
data$Tank <- NULL
data_normalized <- spread(data, Ramp, PAM..Fv.Fm.)
head(data_normalized)
data_normalized$t35_normalizedto28 <- data_normalized$`35` - data_normalized$`28`
data_normalized$t34_normalizedto28 <- data_normalized$`34` - data_normalized$`28`
data_normalized$t33_normalizedto28 <- data_normalized$`33` - data_normalized$`28`
head(data_normalized)

#plots not showing outliers 

#ggplot(data=data_normalized, aes(x=Site, y = t33_normalizedto28, fill = Site)) + geom_boxplot() + theme_minimal() + ylab("PAM Fv/Fm \n 33C normalized to 28C \n (higher = healthier symbionts)") + xlab("") + theme(axis.text=element_text(size=14)) + ylim(-0.15,0.1)

#ggsave("../../CSUMB/Thesis/Data analysis/Figures - bleaching phys/10232020/PAM_lowtemps_normalized.pdf", height = 5, width = 8)

#ggplot(data=data_normalized, aes(x=Site, y = t34_normalizedto28, fill = Site))+ geom_boxplot() + theme_minimal() + ylab("PAM Fv/Fm \n 34C normalized to 28C \n (higher = healthier symbionts)") + xlab("") + theme(axis.text=element_text(size=14)) + ylim(-0.15,0.15)

#ggsave("../../CSUMB/Thesis/Data analysis/Figures - bleaching phys/10232020/PAM_medtemps_normalized.pdf", height = 5, width = 8)

#ggplot(data=data_normalized, aes(x=Site, y = t35_normalizedto28, fill = Site))+ geom_boxplot() + theme_minimal()  + ylab("PAM Fv/Fm \n 35C normalized to 28C \n (higher = healthier symbionts)") + xlab("") + theme(axis.text=element_text(size=14)) + ylim(-0.15,0.15)

#ggsave("../../CSUMB/Thesis/Data analysis/Figures - bleaching phys/10232020/PAM_hightemps_normalized.pdf", height = 5, width = 8)

#all in one plot

data_normalized$StressRecovTimePoint <- data_normalized$Time.point
data_normalized$Time.point <- NULL

data_normalized_gather <- pivot_longer(data_normalized, cols = starts_with("t"), names_to = "temperature_norm", values_to = "pam" )


data_normalized_gather_nas <- data_normalized_gather[!is.na(data_normalized_gather$pam),]
 
  
ggplot(data=data_normalized_gather_nas, aes(x=Site, y = pam, fill = temperature_norm))+ geom_boxplot(outlier.size = -1) + theme_minimal() + scale_fill_manual(name = "Temperature", labels = c("33C", "34C", "35C"), values = c("mediumseagreen", "darkcyan", "darkgoldenrod")) + ylab("Photochemical Efficiency (Fv/Fm)") + xlab("") + theme(axis.text=element_text(size=14)) + ylim(-0.13,0.1)

#ggsave("../../CSUMB/Thesis/Data analysis/Figures - bleaching phys/02152021- normalized to controls/PAM_normalized.pdf", height = 5, width = 8)

```




