---
title: 'Modeling Thermotolerance by Symbionts + pollution 2014 and 2019'
author: "Melissa Naugle"
date: "8/5/2021"
output: html_document
---

```{r setup, include=FALSE}
setwd(dir = "~/Desktop/GitHub/CBASS_bleachingdata/Modeling/")
rm( list = ls())
graphics.off()
library(reshape2)
library(DescTools)
library(epitools)
library(patchwork)
library(tidyverse)
library(mosaic)
library(lme4)
library(lmerTest)
library(mgcv)
```
Load data, join together  

```{r}
LD50_2014 <- read.csv("data/2014_ColonyOutput_Assays_SCLERA.csv")
LD50_2014 <- LD50_2014[LD50_2014$Species == "Acropora hyacinthus",]
LD50_2019 <- read.csv("data/2019_UniqueColony_SCLERAModeling_v1.5.csv")
LD50_2019$Site_Col <- paste0(LD50_2019$Site,"_",LD50_2019$Colony)
LD50_2019$Site <- NULL
LD50_2019$LBSP <- NULL
LD50_2019$pop2016 <- NULL
symbiont_2014 <- read.csv("data/2014_fromtom_symbiont_log_ratios_allsites_08_05_21_LBSP.csv")
symbiont_2014$Site <- NULL
symbiont_2014$Season <- NULL
symbiont_2019 <- read.csv("data/2019_symbiont_log_ratios_08_05_21_LBSP.csv")
symbiont_2019$Site_Col <- paste0(symbiont_2019$Site,"_",symbiont_2019$Colony)

GE_controls_2019 <- read.csv("data/eignegenes_controls.csv")
GE_controls_2019$Site_Col <- paste0(GE_controls_2019$Site,"_",GE_controls_2019$Rep)
GE_controls_2019$Site <- NULL

data_2014 <- left_join(symbiont_2014, LD50_2014, by = "Colony")
data_2019 <- left_join(symbiont_2019, LD50_2019, by = "Site_Col")
data_2019 <- left_join(data_2019, GE_controls_2019, by = "Site_Col")
```

#Modeling 2014

Tried lmer linear model with all factors to describe variation of 1. MortDegDays 2. MortTemp and 3. TimeToMortality
-also looked at impact of interactions (some are signif, though maybe not biologically relevant?)

No models matched the values in the CRCP report, some potential reasons: 
-differences in thermotolerance metric?
-incorrect calculation of C:D ratio (maybe since I capped at 20 and -20 this is causing an issue? though models look similar when using proportion of C)
-something else?

Best model looks like GAM with proportion C, season, and lbsp! (end of chunk)

```{r }

#MortDegDays with full model 
m2014_full_mdd <- lmer(MortDegDays ~ CDratio_log + Season + LBSP + (1|Site), data = data_2014)
summary(m2014_full_mdd)
anova(m2014_full_mdd)
#CD almost signif, LBSP is not

#MortTemp with full model 
m2014_full_mt <- lmer(MortTemp ~ CDratio_log + Season + LBSP + (1|Site), data = data_2014)
summary(m2014_full_mt)
anova(m2014_full_mt)
#LBSP closish to signif, CD is not

#Timeto mortality with full model 
m2014_full_tm <- lmer(TimeToMortality ~ CDratio_log + Season + LBSP + (1|Site), data = data_2014)
summary(m2014_full_tm)
anova(m2014_full_tm)
plot(m2014_full_tm)
#CD is signif, LBSP is not
#closest to ANOVA table from CRCP report 

#proportions C and D to compare to C:D ratio 
m2014_full_mdd_propc <- lmer(MortDegDays ~ C_prop + Season + LBSP + (1|Site), data = data_2014)
anova(m2014_full_mdd_propc)
m2014_full_tm_propc <- lmer(TimeToMortality ~ C_prop + Season + LBSP + (1|Site), data = data_2014)
anova(m2014_full_tm_propc)
#similar to above models using CD ratio. symbionts and season signif, LBSP is not  

AIC(m2014_full_mdd, m2014_full_mt, m2014_full_tm, m2014_full_mdd_propc, m2014_full_tm_propc)
#MortTemp has best fit 

#GAM 

#just using C:D 
m2014_gam_cdonly <- gam(MortDegDays ~ CDratio_log , data =  data_2014)
summary(m2014_gam_cdonly)
#this is signif by symbionts

m2014_gam_full <- gam(MortDegDays ~ CDratio_log + Season + LBSP, data =  data_2014)
summary(m2014_gam_full)
anova(m2014_gam_full)
#signif most by season, then lbsp, close to signif for symbionts 

#not sure about this smoothing part
m2014_gam_full_smoothCD <- gam(MortDegDays ~ s(CDratio_log) + Season + LBSP, data =  data_2014)
summary(m2014_gam_full_smoothCD)
anova(m2014_gam_full_smoothCD)
plot(m2014_gam_full_smoothCD, residuals = T)

######################### Look here 

#s() for continuous variables 
#k = num knots (choose 3 or 5 or something small)
#use gam to see if non-linear then return to lmer w modifications ? 
#could consider using site #GAMM for that - more for exploratory 
#could log transforming C prop - or switch to cd ratio

#try GAM with proportion C
m2014_gam_full_Cprop <- gam(MortDegDays ~ s(log(C_prop+1e-5), k = 5) + Season + LBSP, data =  data_2014)
summary(m2014_gam_full_Cprop)
anova(m2014_gam_full_Cprop)
plot(m2014_gam_full_Cprop, pages = 1, residuals = T)
#everything is significant! 
#linear is better 
#try in lmer 

#try dropping variables and compare fit 
m2014_gam_noCprop <- gam(MortDegDays ~  Season + LBSP, data =  data_2014)
m2014_gam_noseason <- gam(MortDegDays ~  s(C_prop, k = 3)  + LBSP, data =  data_2014)
m2014_gam_nolbsp <- gam(MortDegDays ~  s(C_prop, k = 3) + Season , data =  data_2014)


AIC(m2014_gam_full_Cprop, m2014_gam_noCprop,m2014_gam_noseason, m2014_gam_nolbsp)
#full model is best!! 

plot(sort(data_2014$C_prop))
plot(sort(log10(data_2014$C_prop+1e-5)))

#continue here with lmer similar model to gam 

```

```{r }
#look at % explained by each variable
m2014 <- gam(MortDegDays ~ C_prop, data =  data_2014)
summary(m2014)
#proportion C explained 7.5%

m2014 <- gam(MortDegDays ~ Season, data =  data_2014)
summary(m2014)
#season explains 43.8%

m2014 <- gam(MortDegDays ~ LBSP, data =  data_2014)
summary(m2014)
#LBSP explains 15.4%


m2014 <- gam(MortDegDays ~ Site, data =  data_2014)
summary(m2014)
anova(m2014)
#site explains 55%

#we should probably include site in here somehow?
#better to compare significance 

#################### do this 
#leave one out deviance explained 
#deviance explained with full model and then model - 1 variable  

```


#Modeling 2019

As expected, CD ratio nor pollution is significant 

Same for non-linear- nothing looks to really explain variation 

#### do this 
should maybe go back and try GE data along with c prop and lbsp and ...
include site as random 


```{r }
m2019 <- lmer(Phenotype.X0.mn ~ CDratio_log + LBSP + (1|Site), data = data_2019)
summary(m2019)
anova(m2019)
#neither is signif at all
plot(m2019)

#GAM 

m2019_gam <- gam(Phenotype.X0.mn ~ CDratio_log + LBSP, data =  data_2019)
summary(m2019_gam)
anova(m2019_gam)
#nothing signif 

m2019_gam_cdonly <- gam(Phenotype.X0.mn ~ CDratio_log, data =  data_2019)
summary(m2019_gam_cdonly)
#symbiont not signif

m2019_gam_lbsponly <- gam(Phenotype.X0.mn ~ LBSP, data =  data_2019)
summary(m2019_gam_lbsponly)
anova(m2019_gam_lbsponly)
#lbsp not signif 

m2019_gam_siteonly <- gam(Phenotype.X0.mn ~ Site, data =  data_2019)
summary(m2019_gam_siteonly)
anova(m2019_gam_siteonly)
#site is not signif 

#try proportion C

m2019_gam_all_propC <- gam(Phenotype.X0.mn ~ C_prop + LBSP + Site, data =  data_2019)
anova(m2019_gam_all_propC)
#nothing signif 
m2019_gam_propC_lbsp <- gam(Phenotype.X0.mn ~ C_prop + LBSP , data =  data_2019)
anova(m2019_gam_propC_lbsp)
#nothing signif 

# try GE
m2019 <- lm(Phenotype.X0.mn ~ MEcoral2 + MEthistle1 + MEbisque4 + MEpurple + MEdarkorange + MEpaleturquoise + MEdarkmagenta + MEorange + MEhoneydew1 + MEorangered4, data =  data_2019)
anova(m2019)
summary(m2019)
sm2019 <- step(m2019)
summary(sm2019)
#overfitting an issue 
#n/10 -> should use ~4 variables 
#don't rely on step -> instead use train test (80% to train, then test using remaining 20%)
#get r2 -> randomize and bootstrap 10,000 times and get distribution 

#only select signif modules 

sigm2019 <- lm(Phenotype.X0.mn ~ MEthistle1 + MEpurple + MEpaleturquoise + MEdarkmagenta + MEdarkorange + MEhoneydew1, data =  data_2019)
anova(sigm2019)
summary(sigm2019)

#simplify remove non sig 
sigm2019_simp <- lm(Phenotype.X0.mn ~ MEthistle1 + MEpurple + MEpaleturquoise + MEdarkmagenta, data =  data_2019)
anova(sigm2019_simp)
summary(sigm2019_simp)
#r2 is 66% 

AIC(sigm2019_simp, sigm2019)
#maybe retain all 6 modules?
```

```{r}

```


5.02% of thermotolerance variation explained by proportion C in 2019 
(compared to 7.5% in 2014)

0.0272% of thermotolerance variation explained by LBSP in 2019
(compared to 15.4% in 2014)

Better to do this as 'step-down' method to determine var explained 

```{r }
#look at % var explained by each variable
m2019 <- gam(Phenotype.X0.mn ~ MEthistle1, data =  data_2019)
summary(m2019)
#14.8% by thistle1

m2019 <- gam(Phenotype.X0.mn ~ MEpurple, data =  data_2019)
summary(m2019)
#13.4 by purple

m2019 <- gam(Phenotype.X0.mn ~ MEpaleturquoise, data =  data_2019)
summary(m2019)
#13.3 by paleturquoise

m2019 <- gam(Phenotype.X0.mn ~MEdarkmagenta, data =  data_2019)
summary(m2019)
#22.8 by darkmeganta

m2019 <- gam(Phenotype.X0.mn ~ C_prop, data =  data_2019)
summary(m2019)
#5.02 by C prop

m2019 <- gam(Phenotype.X0.mn ~ LBSP, data =  data_2019)
summary(m2019)
#0.0272 % by LBSP
```



########################################################################################


#Plot symbiont community by thermotolerance (Fig 6a from report)

Weird that 2014 has so many 100% D colonies, but report doesn't show that? 
-fig 6a high thermotol,low C points are mostly high pollution, maybe driven by Cannery data that we do not have here?

Though it still looks like more of a trend in 2014 compared to 2019 ?

```{r}
#2014
g14 <- ggplot(data_2014, aes(x= CDratio_log, y = MortDegDays #,col = Site
                             )) + geom_point(size = 5) + xlab("Log C:D Ratio \n left = more D, right = more C") + ylab("MortDegDays \n higher = more thermotol") + theme_bw() + ggtitle("2014 Log C:D ratio by Mortality in Degree Days") + geom_smooth(method=loess) + xlim(-21,21)
#tried method = gam here and looks almost linear 

#2019
g19 <- ggplot(data_2019, aes(x= CDratio_log, y = Phenotype.X0.mn #,col = Site
                             )) + geom_point(size = 5) + xlab("Log C:D Ratio \n left = more D, right = more C") + ylab("Two-metric thermotolerance \n higher = more thermotol") + theme_bw() + ggtitle("2019 Log C:D ratio by two-metric thermotolerance") + geom_smooth(method= gam)  + xlim(-21,21)
#tried method = loess here and won't plot 

g14+g19

```
Try proportion C / D instead of ratio

```{r}
#2014
g14 <- ggplot(data_2014, aes(x= C_prop, y = MortDegDays #,col = Site
                             )) + geom_point(size = 5) + xlab("Proportion C") + ylab("MortDegDays \n higher = more thermotol") + theme_bw() + ggtitle("2014 Log C:D ratio by Mortality in Degree Days") + geom_smooth(method=gam) 

#2019
g19 <- ggplot(data_2019, aes(x= C_prop, y = Phenotype.X0.mn #,col = Site
                             )) + geom_point(size = 5) + xlab("Log C:D Ratio \n left = more D, right = more C") + ylab("Two-metric thermotolerance \n higher = more thermotol") + theme_bw() + ggtitle("2019 Log C:D ratio by two-metric thermotolerance") + geom_smooth(method=gam) + xlim(0,1)

g14+g19

```


Try removing the 100% C and 100% D colonies to recreate 6a?

Then try using MortTemp instead of MortDegDays

```{r}
#2014

data_2014_test <- data_2014[!data_2014$CDratio_log == -20,]
data_2014_test <- data_2014_test[!data_2014_test$CDratio_log == 20,]
ggplot(data_2014_test, aes(x= CDratio_log, y = MortDegDays)) + geom_point(size = 5) + xlab("Log C:D Ratio \n left = more D, right = more C") + ylab("MortDegDays \n higher = more thermotol") + theme_bw() + ggtitle("2014 Log C:D ratio by Mortality in Degree Days") + stat_smooth(method=gam)

ggplot(data_2014, aes(x= CDratio_log, y = MortTemp)) + geom_point(size = 5) + xlab("Log C:D Ratio \n left = more D, right = more C") + ylab("MortTemp \n higher = more thermotol") + theme_bw() + ggtitle("2014 Log C:D ratio by Mortality Temperature") + stat_smooth(method=gam)

```
