---
title: 'Modeling Thermotolerance by Symbionts + pollution 2014 and 2019'
author: "Melissa Naugle"
date: "8/5/2021"
output: html_document
---

```{r setup, include=FALSE}
setwd(dir = "~/Desktop/GitHub/CBASS_bleachingdata/Modeling/")
rm( list = ls())
graphics.off()
library(reshape2)
library(DescTools)
library(epitools)
library(patchwork)
library(tidyverse)
library(mosaic)
library(lme4)
library(lmerTest)
library(mgcv)
```
Load data, join together  

```{r}
LD50_2014 <- read.csv("data/2014_ColonyOutput_Assays_SCLERA.csv")
LD50_2014 <- LD50_2014[LD50_2014$Species == "Acropora hyacinthus",]
LD50_2019 <- read.csv("data/2019_UniqueColony_SCLERAModeling_v1.5.csv")
LD50_2019$Site_Col <- paste0(LD50_2019$Site,"_",LD50_2019$Colony)
LD50_2019$Site <- NULL
LD50_2019$LBSP <- NULL
LD50_2019$pop2016 <- NULL
symbiont_2014 <- read.csv("data/2014_fromtom_symbiont_log_ratios_allsites_08_05_21_LBSP.csv")
symbiont_2014$Site <- NULL
symbiont_2014$Season <- NULL
symbiont_2019 <- read.csv("data/2019_symbiont_log_ratios_08_05_21_LBSP.csv")
symbiont_2019$Site_Col <- paste0(symbiont_2019$Site,"_",symbiont_2019$Colony)

GE_controls_2019 <- read.csv("data/eignegenes_controls.csv")
GE_controls_2019$Site_Col <- paste0(GE_controls_2019$Site,"_",GE_controls_2019$Rep)
GE_controls_2019$Site <- NULL

data_2014 <- left_join(symbiont_2014, LD50_2014, by = "Colony")
data_2019 <- left_join(symbiont_2019, LD50_2019, by = "Site_Col")
data_2019 <- left_join(data_2019, GE_controls_2019, by = "Site_Col")
```

#Modeling 2014

```{r }
#s() for continuous variables 
#k = num knots (choose 3 or 5 or something small)
#use gam to see if non-linear then return to lmer w modifications ? 
#could consider using site #GAMM for that - more for exploratory (not trying this rn)
#could log transform C prop - or switch to cd ratio

#look at dist of c prop 
plot(sort(data_2014$C_prop))
plot(sort(log10(data_2014$C_prop+1e-5)))

#try GAM full model 
#everything is significant  
m2014_gam_full_cproplog <- gam(MortDegDays ~ s(log(C_prop+1e-5), k = 5) + Season + LBSP + Site, data =  data_2014)
summary(m2014_gam_full_cproplog)

m2014_gam_full_cprop <- gam(MortDegDays ~ s(C_prop, k = 5) + Season + LBSP + Site, data =  data_2014)
summary(m2014_gam_full_cprop)

m2014_gam_full_cdratio <- gam(MortDegDays ~ s(CDratio_log, k = 5) + Season + LBSP + Site, data =  data_2014)
summary(m2014_gam_full_cproplog)

AIC(m2014_gam_full_cproplog, m2014_gam_full_cprop, m2014_gam_full_cdratio)

#cprop is best. not log. 
#use c prop to continue 

#drop variables and compare fit 


#no site
m2014_gam_cprop_sea_lbsp <- gam(MortDegDays ~ s(C_prop, k = 5) + Season + LBSP, data =  data_2014)
summary(m2014_gam_cprop_sea_lbsp)
plot(m2014_gam_full_Cprop, pages = 1, residuals = T)

#no lbsp
m2014_gam_nolbsp <- gam(MortDegDays ~ s(C_prop, k = 5) + Season  + Site, data =  data_2014)
summary(m2014_gam_nolbsp)

#no season
m2014_gam_nosea <- gam(MortDegDays ~ s(C_prop, k = 5) + LBSP  + Site, data =  data_2014)
summary(m2014_gam_nosea)

#no c prop 
m2014_gam_nocprop <- gam(MortDegDays ~ Season +  LBSP  + Site, data =  data_2014)
summary(m2014_gam_nocprop)

AIC(m2014_gam_full_cprop,m2014_gam_cprop_sea_lbsp, m2014_gam_nolbsp,m2014_gam_nosea, m2014_gam_nocprop)
#full model and no lbsp are tied 

#try the same below without site since we will use as random factor in linear model 
#leave-one-out deviance explained - deviance explained with full model and then model - 1 variable  
m2014_gam_full <- gam(MortDegDays ~ s(C_prop, k = 5) + Season + LBSP, data =  data_2014)
summary(m2014_gam_full) #58.5%

m2014_gam_full_nocprop <- gam(MortDegDays ~ Season + LBSP, data =  data_2014)
summary(m2014_gam_full_nocprop) #55%
#full model - c prop: 58.5 - 55 = 3.5% explained by c prop 

m2014_gam_full_noseason <- gam(MortDegDays ~  s(C_prop, k = 5)  + LBSP, data =  data_2014)
summary(m2014_gam_full_noseason) #24.4%
#full model - season: 58.5 - 24.4 = 34.1% explained by season

m2014_gam_full_nolbsp <- gam(MortDegDays ~  s(C_prop, k = 5)  + Season, data =  data_2014)
summary(m2014_gam_full_nolbsp) #47.1%
#full model - lbsp: 58.5 - 47.1 = 11.4% explained by lbsp 

AIC(m2014_gam_full, m2014_gam_full_nocprop,m2014_gam_full_noseason,m2014_gam_full_nolbsp)
#full model is best 

#continue here with lmer similar model to gam 
m2014_lm_full <- lmer(MortDegDays ~ C_prop + Season + LBSP +  (1|Site), data =  data_2014)
summary(m2014_lm_full)

m2014_lm_full_nosite <- lm(MortDegDays ~ C_prop + Season + LBSP, data =  data_2014)
summary(m2014_lm_full_nosite)

AIC(m2014_gam_full, m2014_lm_full, m2014_lm_full_nosite)
#lm full model is best with site 

#use this model!
summary(m2014_lm_full)
anova(m2014_lm_full)

```

#Modeling 2019

Notes:
overfitting an issue? rule: n/10 -> so should use ~4 variables since n = 38 
don't rely on step -> instead use train test (80% to train, then test using remaining 20%)
get r2 -> randomize and bootstrap 10,000 times and get distribution 

```{r }
#GAM first 

m2019_gam_all <- gam(Phenotype.X0.mn ~ s(C_prop, k = 5) + LBSP +  MEthistle1 + MEpurple + MEpaleturquoise + MEdarkmagenta + MEdarkorange + MEhoneydew1, data =  data_2019)
summary(m2019_gam_all)
anova(m2019_gam_all)
#remove non-signif lbsp, c prop, honeydew, darkorange 

#added smoothing function since eigengene expression is continuous 
m2019_gam_4mod <- gam(Phenotype.X0.mn ~ s(MEthistle1, k = 5) + s(MEpurple, k = 5) + s(MEpaleturquoise, k = 5) + s(MEdarkmagenta, k = 5), data =  data_2019)
summary(m2019_gam_4mod)
anova(m2019_gam_4mod)
#all 4 signif 

#remove one at a time 
m2019_gam_nothist <- gam(Phenotype.X0.mn ~ s(MEpurple, k = 5) + s(MEpaleturquoise, k = 5) + s(MEdarkmagenta, k = 5), data =  data_2019)
m2019_gam_nopurp <- gam(Phenotype.X0.mn ~s(MEthistle1, k = 5)  + s(MEpaleturquoise, k = 5) + s(MEdarkmagenta, k = 5), data =  data_2019)
m2019_gam_notur <- gam(Phenotype.X0.mn ~s(MEthistle1, k = 5) + s(MEpurple, k = 5)  + s(MEdarkmagenta, k = 5), data =  data_2019)
m2019_gam_nomag <- gam(Phenotype.X0.mn ~s(MEthistle1, k = 5) + s(MEpurple, k = 5) + s(MEpaleturquoise, k = 5) , data =  data_2019)

AIC(m2019_gam_4mod, m2019_gam_nothist,m2019_gam_nopurp,m2019_gam_notur, m2019_gam_nomag)
#best includes all 4 modules 

#now lmer

m2019_lm_4mod <- lmer(Phenotype.X0.mn ~MEthistle1 + MEpurple + MEpaleturquoise + MEdarkmagenta +  (1|Site), data =  data_2019)
summary(m2019_lm_4mod)
anova(m2019_lm_4mod)

m2019_lm_4mod_nosite <- lm(Phenotype.X0.mn ~MEthistle1 + MEpurple + MEpaleturquoise + MEdarkmagenta , data =  data_2019)
summary(m2019_lm_4mod_nosite)
anova(m2019_lm_4mod_nosite)

AIC(m2019_lm_4mod, m2019_lm_4mod_nosite)
#site marginally helps fit 

#final model
summary(m2019_lm_4mod)
anova(m2019_lm_4mod)
```
Train test model for 2019 GE data 

```{r}
#Generate a relatively small data set N~100
N=200
fake=data.frame(x=rnorm(N))
fake$y=1*fake$x+0+.5*(rnorm(N))
fake$e1=rnorm(N)
fake$e2=rnorm(N)
fake$e3=rnorm(N)
fake$e4=rnorm(N)
fake$e5=rnorm(N)
fake$e6=rnorm(N)
fake$e7=rnorm(N)
fake$e8=rnorm(N)
fake$e9=rnorm(N)

#Train on 80-90% of data
Ptrain=.8
#Perform Nruns random draws/model fits
Nruns=100
Ntrain=round(N*Ptrain,0)

#Fit full model with 9 totally random noise parameters (don't actually use these in your model, this is just to show you overfitting)
fullmod=lm(y~x+e1+e2+e3+e4+e5+e6+e7+e8+e9,data=fake)
sfullmod=summary(fullmod)
sfullmod
sfullmod$adj.r.squared

#Loop it Nruns times
TrainOut=data.frame(Run=1:Nruns,Fullmod.rsq=sfullmod$adj.r.squared,trainmod.rsq=NA,test.rsq=NA)
for(i in 1:Nruns){
  #get rows for each set, train and test
  train_i=sample(1:N,Ntrain)
  test_i=setdiff(1:N,train_i)
  
  #get datasets
  train=fake[train_i,]
  test=fake[test_i,]

  #Train the model
  trainmod=lm(y~x+e1+e2+e3+e4,data=train)
  strainmod=summary(trainmod)
  #output train data
  TrainOut$trainmod.rsq[i]=strainmod$adj.r.squared
  
  #test predictions using data the model has yet to see
  ttmod=lm(test$y~predict(trainmod,newdata=test))
  sttmod=summary(ttmod)
  #write results out
  TrainOut$test.rsq[i]=sttmod$adj.r.squared
  #print progress
  print(paste0(i," of Nruns"))
}

par(mfrow=c(3,1))
#plot1
plot(test$y~predict(trainmod,newdata=test))
abline(ttmod)

#plot2:red is full/overfit model fit; Green is train/test
plot(sort(TrainOut$test.rsq),col="gray",ylim=c(0.5,1))
abline(h=median(TrainOut$test.rsq),col="green")
abline(h=median(TrainOut$Fullmod.rsq),col="red")

#plot3: red is full/overfit model fit; Green is train/test
hist((TrainOut$test.rsq),N/10,xlim=c(0.5,1))
abline(v=median(TrainOut$test.rsq),col="green")
abline(v=median(TrainOut$Fullmod.rsq),col="red")

```


########################################################################################


#Plot symbiont community by thermotolerance (Fig 6a from report)

Weird that 2014 has so many 100% D colonies, but report doesn't show that? 
-fig 6a high thermotol,low C points are mostly high pollution, maybe driven by Cannery data that we do not have here?

Though it still looks like more of a trend in 2014 compared to 2019 ?

```{r}
#2014
g14 <- ggplot(data_2014, aes(x= CDratio_log, y = MortDegDays #,col = Site
                             )) + geom_point(size = 5) + xlab("Log C:D Ratio \n left = more D, right = more C") + ylab("MortDegDays \n higher = more thermotol") + theme_bw() + ggtitle("2014 Log C:D ratio by Mortality in Degree Days") + geom_smooth(method=loess) + xlim(-21,21)
#tried method = gam here and looks almost linear 

#2019
g19 <- ggplot(data_2019, aes(x= CDratio_log, y = Phenotype.X0.mn #,col = Site
                             )) + geom_point(size = 5) + xlab("Log C:D Ratio \n left = more D, right = more C") + ylab("Two-metric thermotolerance \n higher = more thermotol") + theme_bw() + ggtitle("2019 Log C:D ratio by two-metric thermotolerance") + geom_smooth(method= gam)  + xlim(-21,21)
#tried method = loess here and won't plot 

g14+g19

```
Try proportion C / D instead of ratio

```{r}
#2014
g14 <- ggplot(data_2014, aes(x= C_prop, y = MortDegDays #,col = Site
                             )) + geom_point(size = 5) + xlab("Proportion C") + ylab("MortDegDays \n higher = more thermotol") + theme_bw() + ggtitle("2014 Log C:D ratio by Mortality in Degree Days") + geom_smooth(method=gam) 

#2019
g19 <- ggplot(data_2019, aes(x= C_prop, y = Phenotype.X0.mn #,col = Site
                             )) + geom_point(size = 5) + xlab("Log C:D Ratio \n left = more D, right = more C") + ylab("Two-metric thermotolerance \n higher = more thermotol") + theme_bw() + ggtitle("2019 Log C:D ratio by two-metric thermotolerance") + geom_smooth(method=gam) + xlim(0,1)

g14+g19

```


Try removing the 100% C and 100% D colonies to recreate 6a?

Then try using MortTemp instead of MortDegDays

```{r}
#2014

data_2014_test <- data_2014[!data_2014$CDratio_log == -20,]
data_2014_test <- data_2014_test[!data_2014_test$CDratio_log == 20,]
ggplot(data_2014_test, aes(x= CDratio_log, y = MortDegDays)) + geom_point(size = 5) + xlab("Log C:D Ratio \n left = more D, right = more C") + ylab("MortDegDays \n higher = more thermotol") + theme_bw() + ggtitle("2014 Log C:D ratio by Mortality in Degree Days") + stat_smooth(method=gam)

ggplot(data_2014, aes(x= CDratio_log, y = MortTemp)) + geom_point(size = 5) + xlab("Log C:D Ratio \n left = more D, right = more C") + ylab("MortTemp \n higher = more thermotol") + theme_bw() + ggtitle("2014 Log C:D ratio by Mortality Temperature") + stat_smooth(method=gam)

```
